---
title: "R Basics"
author: "Catherine Yu"
date: "10/24/2016"
output: html_document
---

# 1.1 Variable Assignment
Which to use? `<-` vs. `=`

Answer: _Always_ use `<-`
```{r arrow or equal}
# GOOD
x <- 5

# BAD
x = 5
```

Also _always_ surround your operations and assignments with spaces.
```{r spaces, results = "hide"}
# WHAT IS THIS?
x<-3

# DOES IT MEAN...
x <- 3 

# OR DOES IT MEAN...
x < -3
```

# 1.2 Basic Operations
```{r basic operations}
a <- c(1, 2, 3, 4)
a + 5
a - 10
a * 2
a / 5
```

```{r more complex operations}
sqrt(a)
exp(a)
log(a)
exp(log(a))
```

# 1.3 Index
*R uses 1-based indexing
*Negative indices returns all values EXCEPT at those positions
```{r index}
index_vec <- c("azuki", "cat", "dog")
index_vec[1]
index_vec[c(-2, -3)]
```

# 1.4 Data Types & Data Structures

Data Types:

type     | output            |
---------|-------------------|
logical  | TRUE(1)/FALSE(0)  |
numeric  | 12.3, 5, 999      |
integer  | 2L, 34L, 0L       |
character| "good", "bad", "2"|
factor   | "F", "M"          |

Data Structures:

type | Homogenous | Heterogenous|
-----|------------|-------------|
 1D  | Vector     | Lists       |
 2D  | Matrix     | Data Frames |
 nd  | Array      |             |

Vector `c()` and Lists `list()`
```{r 1D data structures}
b <- c(4:8, a)
class(b)
is.numeric(b)
is.integer(b)
is.character(b)

c <- c("John", "Jane")
class(c)
is.character(c)

d <- c(b, c)
class(d)
d

d <- list(b, c)
d
class(d)
is.list(d)

class(d[[1]])
class(d[[2]])
```

``` {r 1D tips}
# Labeling each element in a vector
decile_values <- 1:10
names(decile_values) <- c("10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")
decile_values

# Can call vector value by the label
decile_values["10%"]

# Slice and dice
which(decile_values > 5)
which.min(decile_values)
min(decile_values)

which.max(decile_values)
max(decile_values)

length(decile_values)
sum(decile_values > 6)

# Vectors of repetitive values
rep(1:2, each = 3)
rep(1:2, times = 3)
```

Matrices and Data Frames
```{r 2D data structures}
# By default, fills in column-wise
matrix(1:12, nrow = 4,  
       dimnames = list(
         c("one", "two", "three", "four"),
         c("a", "b", "c")))

# Can use `byrow` argument to fill int row-wise
matrix(1:12, nrow = 4, byrow = TRUE, 
       dimnames = list(
         c("one", "two", "three", "four"),
         c("a", "b", "c")))

# Data Frames
ex_df <- data.frame("names"     = c("azuki", "cat", "jen", "kevin"),
                    "timestamp" = rep(format(Sys.time()), times = 4),
                    "strings"   = c("HELLO", "PLZ", "SEND", "HALPZ"),
                    "numbers"   = 9:12,
                    stringsAsFactors = F)
ex_df
class(ex_df)
str(ex_df)
```

```{r data frame tips}
names(ex_df)
dim(ex_df)

# Add new column
ex_df$books <- c("arabian nights", "huckleberry finn", "alice in wonderland", "pride and prejudice")
ex_df

# Append a new row
ex_df <- rbind(ex_df, data.frame("names"     = "nicky",
                                 "timestamp" = format(Sys.time()),
                                 "strings"   = "IMMEDIATELY",
                                 "numbers"   = 13,
                                 "books"     = "gone girl"))

# Delete column
ex_df$books <- NULL
ex_df

# Delete row
ex_df[-nrow(ex_df),]
ex_df

ex_df <- ex_df[-nrow(ex_df),]
ex_df
```

# 1.5 How to ask R for help?
```{r help, results = "hide"}
# ? gets you to the documentation for specific functions/calls
?data.frame
?c
?rep

# ?? when you have no idea what functions/calls/library you need
??dataframe
??regression
```

# 1.6 Packages and Libraries
```{r pkg, eval = FALSE}
# If package is on CRAN
install.packages("swirl")
install.packages("data.table")

# Loading libraries
library(swirl)
library(data.table)
```

# 1.7 Working Directory & Environment
```{r wd, results = "hide"}
# Current working directory
getwd()

# Change working directory
setwd("~/Source/r-seminar/")
getwd()

# All objects in the environment
ls()
```

# 1.8 Reading in data from database
```{r, results = "hide"}
library(RMySQL)
library(zdsdb)

# 0 - connect to db
conn_client <- connectToDb(dbhost = "di-abvrds1.zephyr-intranet.com",
                           dbschema = "zdb_abv")
```
```{r reading data from db}

# 1 - get data
query <- "SELECT * FROM ProviderTopicSummaries PTS WHERE PTS.topicId = 68;"
results <- suppressWarnings(dbGetQuery(conn_client, query))

# 2 - what do we have?
dim(results)
str(results)
names(results)

length(unique(results$entityId))
sum(duplicated(results$entityId))
unique(results$topicId)
summary(results$defaultScore)
table(results$guidelinesCount)
table(results$leadershipCount, results$guidelinesCount)

results <- results[order(-results$defaultScore),]
head(results)
```

# 1.8 Reading in data sets
[Decapitated Animals Dataset](https://github.com/uselesspress/decapitated-animals-data)
```{r reading csv, eval = FALSE}
dat <- read.csv("../data/raw/01_animals.csv", stringsAsFactors = F)
write.csv(dat, "../data/processed/01_animals_output.csv", row.names = F)
```
See `01_decap_analysis.R` for sample analysis

